---
title: 'Worksheet 1: R Basics'
author: "David Askew"
date: "1-31-2024"
---
_This is the first in a series of worksheets for History 8510 at Clemson University. The goal of these worksheets is simple: practice, practice, practice. The worksheet introduces concepts and techniques and includes prompts for you to practice in this interactive document. When you are finished, you should change the author name (above), knit your document to a pdf, and upload it to canvas._

## What is R? 
To start let's define what exactly R is. R is a language and environment for statistical computing and graphics. R provides a variety of statistical and graphical techniques and its very extensible which makes it an ideal language for historians.  

## Foundational Concepts
### Values

There are several kinds of variables in R. Numeric, logical, and strings. 

R takes inputs and returns an output. So for example, if our input is a number the output will be a number. Simply typing a number, below or in the console, will return a number or a **numeric** value. 
```{r}
5
```
Same thing happens if we use a number with a decimal. 
```{r}
5.5483
```
Numbers can be used to do arithmetic. 
```{r}
5 + 5
```

(@) You try, multiple two numbers. 
```{r}
5*5
```


(@) Can you multiply two numbers and then divide by the result?
```{r}
10*10/5
```


The next type of value is a **string**. Strings are lines of text. Sometimes these are referred to as character vectors. To create a string, you add text between two quotation marks. For example: 
```{r}
"Go Tigers"
```

(@) Try to create your own string.
```{r}
"David is awesome"
```

You can't add strings using `+` like you can numbers. But there is a function called `paste()` which concatenates character vectors into one. This function is _very_ useful and one you'll use a lot in a variety of circumstances. Here's what that looks like: 

```{r}
paste("Hello", "Clemson Graduate Students")
```

(@) Try it, add two strings together like the above example. 
```{r}
paste("iHola!", "Estudiantes de Universidad de Clemson")
```

(@) Can you explain what happened in 2-3 sentences? 

```{r}
By using paste, I gave R a command to display characters in a message box. Then I added my message in the quotation marks which is what will be displayed and the comma with a new quotation will display another message in the same coding chunk. Therefore, this is  two strings with one  message. For R, this also appears to be the equivalent of using x=MsgBox (0+64) when creating an error box in Windows Notepad. However, R is most likely different by which it does not display an error message on your screen but rather assign a value to a message you create. 
```


The last type are **logical** values like `TRUE` and `FALSE`. (Note that these are all caps.)

```{r}
TRUE
```

```{r}
FALSE
```

These logical values are really useful for testing a statement and comparing values to each other. For example: 
```{r}
3 < 4
```
3 is indeed less than 4 so the return value is `TRUE`. Here are a few more examples:

R also has relational operators and logical operators. Relational operators test how one object relates to another. We've been using some of these above. 

* Equality `==`
* Inequality `!=`
* Less than or Greater than `<` and `>`
* Less than/greater than or equal to `<=` and `>=`

Logical Operators allow you to combine or change the results of comparisons. 
* AND `&`
* OR `|`
* NOT `!`

```{r}
5 == 10
3 < 4
5 != 10
5 != 5
3 == 4 | 3
3 != 4 & 3 != 5
```
(@) Explain, what does the code on each line above do? 

>
```{r}
[1] TRUE
> 5 == 10
[1] FALSE
> 5 != 5
[1] FALSE
> 3 == 4 | 3
[1] TRUE
> 3 != 4 & 3 != 5
[1] TRUE

These lines perform actions that help us explain differences and similarities in how various operators function and how they can be interpreted.
```

(@) Create your own comparison. 
```{r}
[1] TRUE
> 7==7
[1] TRUE
> 6!=6
[1] FALSE
> 2*4==8
[1] TRUE
> 4*10!=40
[1] FALSE
> 6*5/2==14
[1] FALSE
```

---

### Variables

Values are great but they are made so much more powerful when combined with **Variables** which are a crucial building block of any programming language. Simply put, a variable stores a value. For example if I want x to equal 5 I can do that like this: 
```{r}
x <- 5
```
`<-` is known as an assignment operator. Technically, you could use `=:` here too but it is considered bad practice and can cause complicated issues when you write more advanced code. So its important to stick with `<-` whenever you are coding in R. 

I can also add a string or character vector to a variable. 
```{r}
x <- "Go Tigers!"
```

Variable names can be almost anything. 
```{r}
MyFavoriteNumber <- 25
```
Variable or object names must start with a letter, and can only contain letters, numbers, `_`, and `.`. You want your object names to be descriptive, so you’ll need a convention for multiple words. People do it different ways. I tend to use periods but there are several options: 

```
i_use_snake_case 
otherPeopleUseCamelCase 
some.people.use.periods 
And_aFew.People_RENOUNCEconvention
```

Whatever you prefer, be consistent and your future self will thank you when your code gets more complex. 

(@) You try, create a variable and assign it a number: 

```{r}
variable <- 56
```

(@) Can you assign a string to a variable?

```{r}
createstring <- "I like coffee"
```

Once we've assigned a variable we can use that variable to run calculations just like we did with raw numbers.
```{r}
x <- 25
x * 5
```

R can also handle more complex equations. 
```{r}
(x + x)/10
(x + x * x) - 100
```

And we could store the output of a calculation in a new variable:
```{r}
My.Calculation <- (x + x)*10
```

(@) You try. Assign a number to x and a number to y. Add those two numbers together.
```{r}
> x <- 12
> y <- 64
> x + y
[1] 76
```

(@) Can you take `x` and `y` and multiply the result by 5? 
```{r}
> x + y * 5
[1] 332
```

(@) Try creating two variables with names other than `x ` and `y`. Descriptive names tend to be more useful. Can you multiply the contents of your variables? 
```{r}
> empanadas * jamon
[1] 63
```

(@) Try creating two variables that store strings. Can you concatenate those two variables together? 
```{r}
> a <- "Mater's"
> b <- "student"
> paste(a, b, sep="")
[1] "Mater'sstudent"
> paste(a, b, sep=" ")
[1] "Mater's student"
```

### Vectors

If we have a lot of code and rely on just variables, we're going to have a lot of variables. That's where **vectors** come into play. Vectors allow you to store multiple values. All variables in R are actually already vectors. That's why when R prints an output, there is a `[1]` before it. That means there is one item in that vector. 

```{r}
myvalue <- "George Washington"
myvalue
```
In this instance "George Washington" is the only item in the variable myvalue. But we could add more. To do that we use the `c()` function which combines values into a vector. 
```{r}
myvalue <- c("George Washington", "Franklin Roosevelt", "John Adams")
myvalue
```
You'll notice that the output still only shows `[1]` but that doesn't mean there is only one item in the list. It simply means George Washington is the first. If we use `length()` we can determine the number of items in this vector list. 
```{r}
length(myvalue)
```
We could get the value of the 2nd or 3rd item in that list like this: 
```{r}
myvalue[2]
```
We could also create a vector of numbers:
```{r}
my.numbers <- c(2, 4, 6, 8)
```

And we could then do calculations based on these values: 
```{r}
my.numbers * 2
```
(@) Explain in a few sentences, what happened in the code above?
```{r}
> numericvalues <- c(1, 3, 5, 7, 9)
> my.numbers <- c(2, 4, 6, 8)
> my.numbers * 2
[1]  4  8 12 16

What happened was that I assigned numbers to represent the value numericvalues and did the same for my.numbers. After doing this for both values, I multiplied by 2 to find the existing code demonstrated in the values of 4, 8, 12, and 16. 
```

>

Lets try something slightly different. 
```{r}
my.numbers[3] * 2
```
(@) Explain in a few sentences, what happened in the code above? 

>
```{r}
> my.numbers [3]*2
[1] 12

My numbers were 2, 4, 6, 8 and the third number was 6. So when I multiplied 6 * 2, the value came to equal 12. This is because the bracket indicates a third symbol of the code. 
```

(@) You try, create a list of five items and store it in a descriptive variable. 
```{r}
FavoriteShows <- list("Lost", "Game of Thrones", "Heroes", "Adventure Time", "Rick and Morty")
```


### Built In Functions

R also has **built in functions**. We've already used a couple of these: `paste()` and `c()`. But there are others, like `sqrt()` which does what you think it does, finds the square root of a number. 
```{r}
sqrt(1000)
```

Many functions have options that can be added to them. For example, the `round()` function allows you to include an option specifying how many digits to round to. 

You can run it without that option and it'll use the default: 
```{r}
round(15.492827349)
```

Or we can tell it to round it to 2 decimal places.
```{r}
round(15.492827349, digits = 2)
```

How would you know what options are available for each function in R? Every function and package in R comes with **documentation** or a **manual** that is built into R studio and can be pulled by by typing a question mark in front of the function in your console. These packages will commonly give you examples of how to use the function and syntax for doing so. They are incredibly useful. 

We can ask R to pull up the documentation like this:
```{r}
?round()
```

(@) Now you try, find the documentation for the function  `signif()`.
```{r}
> signif(3.141593, digits=7)
[1] 3.141593
> signif(3.141593, digits=2)
[1] 3.1
```

In real life, you typically you wouldn't want to store this code in your script file. You probably don't need to pull up the documentation for the function every time you run that piece of code. But for the purposes of this worksheet we're adding it to our `.Rmd` document. 

(@) Use the console to find the documentation for `floor()`? Try it and then tell me, what does that function do? 

>
```{r}
> floor (3.1)
[1] 3
> floor (8.9999)
[1] 8

What this function does is round the number to the closest value, but for some numbers it does not perform this task like with 8.9999. I am not sure why it did this, but it might be because it  goes to the smallest value instead of the next number. 
```

### Data Frames & Packages

R is the language of choice for most data scientists and that is because of its powerful suite of data analysis tools. Some are built into R, like `floor()` which we looked at above. But others come from packages that you have to install. Most programming languages have some sort of package system although every language calls it a slightly different thing. Ruby has gems, python has eggs, and php has libraries. In R these are called packages or libraries and they are hosted by the Comprehensive R Archive Network or more commonly, CRAN. CRAN is a network of ftp and web servers around the world that store identical, up-to-date, versions of code and documentation for R. You didn't know it but you already used CRAN when you looked up the documentation for `floor()` above. If you go to the CRAN webpage and look at the [list of available R packages](https://cran.r-project.org/web/packages/available_packages_by_name.html) you'll see just how many there are!

So there are many packages but there are also some that are indispensable. You'll use some packages over and over for this class. We'll get into some of those in the next two worksheets but for now lets look at one basic package for data. 

Lets start with the `tibble()` package which allows you to create and work with **data frames**. What is a data frame? Think of it as a spreadsheet. Each column can contain data - including numbers, strings, and logical values. 

Lets load the `tibble()` package. If you have this package installed this line of code will work. If not, you'll get an error that says something like: `## Error in library("tibble"): there is no package called 'tibble'`. If that's the case (it probably is), then no worries - we can install it. To install a package run `install.packages("tibble")` in your console and R will download and install the package. (Remember this isn't code that we want or need to store in our document, so run it in your console not in the `.Rmd` document.)
```{r}
library(tibble)
```

The tibble package will allow you to create dataframes. Lets use some built in data in R to demonstrate this. This dataset contains the measurements in centimeters of sepal and petal width and length. Lets start by figuring out what kind of data iris is: 
```{r}
typeof(iris)
```

We can see that iris is a list. But we want to use the tibble package to turn it into a dataframe. Thats very easy, we can simply run: 
```{r}
as_tibble(iris)
```

That is useful. But we can also create our own data frame from vectors. That means we can create the values in each field. For example: 

```{r}
tibble(
x = 1:5, 
y = 1, 
z = x * 2 + y
)
```
What is going on here? Each line within this function (x,y,z) creates a new column and determines the values in each row. So the first line creates five rows with values sequentially from 1 to 5. The second column contains only the value 1 in all five rows. 

(@) What is going on in row z? Can you explain the logic here?

>
```{r}
library(tibble)
> tibble(
+     x = 1:5, 
+     y = 1, 
+     z = x * 2 + y
+ )
# A tibble: 5 × 3
      x     y     z
  <int> <dbl> <dbl>
1     1     1     3
2     2     1     5
3     3     1     7
4     4     1     9
5     5     1    11

What is happening in row z is that since y remains the same that equals 1, and x can be anywhere from 1 to 5, the value for z could range into the following - 3, 5, 7, 9, 11 which I could find with the help of the above code.


```


## Loading Dataframes from Packages

Lets also install the data package that I've created for our class. This package contains a variety of historical datasets that you can use to complete your assignments this semester. However, this package is not on CRAN. That's okay. We can still install it from GitHub. To do that we'll need to use the `devtools` package. We'll install devtools and then use it to install our class's package which is hosted on GitHub. 

First, install the `devtools` package.

Our class's package is called `DigitalMethodsData` and its hosted on [GitHub](https://github.com/regan008/DigitalMethodsData). Use the directions on the repository page and what you learned about packages above to install this package. 

After installing the package, we should load it: 
```{r}
library(DigitalMethodsData)

```

What kind of datasets are available in this package? We can use the help documentation to find out. 

Run `help(package="DigitalMethodsData")` in your console to pull up a list of datasets included in this package. 

For demonstration purposes we'll use the `gayguides` data here. Pull up the help documentation for this dataset. What does the documentation tell you abou the scope of this dataset? 

>
```{r}
The documentation discusses Bob Damron's Address Books and includes information transcribed from 1965 to 1985. However, it was last updated on December 2022 and the documentation demonstrates the scope of the dataset to be something which includes data of over the past 20 years between 1965 and 1985. This is handy because it allows the user to access certain observations of data efficiently in an entire set which contains 60,000. Furthermore, it lists the different addresses, type of amenity, city, state, and year these LGBTQ individuals visited and even provides the coordinates. 
```

To use a dataset included in this package we first need to load it. We can do so like this:

```{r}
data(gayguides)
```

Notice that you now have a loaded dataset in your environment pane. It shows us that there are 60,698 observations (rows) of data and 14 variables (columns).

Let's now look at our data. You can use the `head()` function to return the first part of an object or dataset.
```{r}
head(gayguides)
```
That gives us the first 6 rows of data. Its really useful if you just want to peak into the dataset but don't want to print out all 60k+ rows.

(@) The default is to print six rows of data. Can you modify the above code to print out the first 10 rows? 
```{r}
> head(gayguides, n=10L)
    X   ID        title                        description streetaddress           type amenityfeatures           city
1   1 3213       'B.A.'                    (woods & ponds)               Cruising Areas     Cruisy Area    Lake Placid
2   2 2265 'B.A.' Beach                           2 mi. E.        Rte. 2 Cruising Areas     Cruisy Area           Troy
3   3 3269 'B.A.' Beach                 nr. Salt Air Beach               Cruising Areas     Cruisy Area Salt Lake City
4   4 3388 'B.A.' Beach      nr. Evergreen Floating Bridge               Cruising Areas     Cruisy Area        Seattle
5   5 3508 'B.A.' Beach               (2 mi. E. on Rte. 2)               Cruising Areas     Cruisy Area           Troy
6   6 5116 'B.A.' Beach      nr. Evergreen Floating Bridge               Cruising Areas     Cruisy Area        Seattle
7   7  780 'B.A.' Beach on Russian River, at Wohler Bridge               Cruising Areas     Cruisy Area  Russian River
8   8 3810 'B.A.' Beach                           2 mi. E.        Rte. 2 Cruising Areas     Cruisy Area           Troy
9   9 4979 'B.A.' Beach                 nr. Salt Air Beach               Cruising Areas     Cruisy Area Salt Lake City
10 10 5191 'B.A.' Beach      nr. Evergreen Floating Bridge               Cruising Areas     Cruisy Area        Seattle
   state Year notes      lat        lon                                                                     status
1     NY 1982       44.27949  -73.97987 Location could not be verified. General city or location coordinates used.
2     NY 1981       42.72841  -73.69178 Location could not be verified. General city or location coordinates used.
3     UT 1981       40.74781 -112.18727 Location could not be verified. General city or location coordinates used.
4     WA 1981       47.60621 -122.33207 Location could not be verified. General city or location coordinates used.
5     NY 1982       42.72841  -73.69178 Location could not be verified. General city or location coordinates used.
6     WA 1983       47.60621 -122.33207 Location could not be verified. General city or location coordinates used.
7     CA 1983       38.91163 -123.01105 Location could not be verified. General city or location coordinates used.
8     NY 1983       42.72841  -73.69178 Location could not be verified. General city or location coordinates used.
9     UT 1983       40.74781 -112.18727 Location could not be verified. General city or location coordinates used.
10    WA 1984       47.60621 -122.33207 Location could not be verified. General city or location coordinates used.
```

(@) Can you find the last 6 rows of data? 
```{r}
> tail(gayguides)
          X    ID           title               description streetaddress       type
60693 60693 PR351       Boccachio            (Natives only)               Bars/Clubs
60694 60694 PR366       Boccachio            (Natives only)               Bars/Clubs
60695 60695 PR377       Boccachio            (Natives only)               Bars/Clubs
60696 60696 GU002   Mi Elena Club (Inquire locally) (Agana)          <NA> Bars/Clubs
60697 60697 GU015 Star Cafe & Bar Downtown, opp. Agana Loop          <NA> Bars/Clubs
60698 60698 GU019 Star Cafe & Bar Downtown, opp. Agana Loop          <NA> Bars/Clubs
                                                                                  amenityfeatures     city state Year notes
60693                                                                                     (G),(*) Hato Ray    PR 1977      
60694                                                                                         (G) Hato Ray    PR 1979      
60695                                                                                         (L) Hato Ray    PR 1980      
60696                                                         (M) - Mixed Crowds - Some Straights    Agana    GU 1980  <NA>
60697 (M) - Mixed Crowds - Some Straights,(YC) - Young and/or collegiate types,(*) - Very popular    Agana    GU 1975  <NA>
60698 (M) - Mixed Crowds - Some Straights,(YC) - Young and/or collegiate types,(*) - Very popular    Agana    GU 1974  <NA>
           lat       lon                                                                     status
60693 18.42258 -66.05096 Location could not be verified. General city or location coordinates used.
60694 18.42258 -66.05096 Location could not be verified. General city or location coordinates used.
60695 18.42258 -66.05096 Location could not be verified. General city or location coordinates used.
60696 13.47628 144.75022 Location could not be verified. General city or location coordinates used.
60697 13.47628 144.75022 Location could not be verified. General city or location coordinates used.
60698 13.47628 144.75022 Location could not be verified. General city or location coordinates used.
```

(@) Reflect on the previous two prompts. How did you figure these out? 

>
```{r}
I figured it out by using ? for help which brought me to the instructions on how to return the first and last parts of an object. Since head(gayguides) represents the first ten lines, an individual can use head(gayguides), then insert the number of observations they would like to access with n= _L. N represents the number of lines the user wants displayed. For displaying the last the last six lines, tail represents the final observations in the entire data set. Therefore, the user must type tail(gayguides) into the console to see those final observations. 
```

The `str()` function is another very useful function for understanding a dataset. It compactly displays the internal structure of an R object. 
```{r}
str(gayguides)
```
There is a ton of useful info here. First, we see that this is a data.frame and that it has 60698 obs. of  14 variables. Second, we see the names of all the variables (columns) in the dataset. Lastly, it shows us what type of data is contained in each variable. For example, the variable `state` is a chr or character vector while `Year` (note the capitalization) is numeric. 

The `$` operator allows us to get a segment of the data. 
```{r}
head(gayguides$title)
```
While that's useful, this particular dataset happens to be sorted by title. So the first rows are all labeled some version of B.A. beach. That's not very useful. What if we want to see the 500th item in this list? Well, we can pull that up like this: 
```{r}
gayguides$title[500]
```

(@) Break this down. Whats going on here? What do each of the elements of this code mean? (gayguides, $, title, and [500]) 

>
```{r}
(Gayguides) represents the data set of all LGBTQ individuals, $ is a subset for representing the occupational status of each individual regarding employment, title is a function that can be used to add labels to a plot which would include names, calls, or expressions, and [500] is a number which symbolizes the 500th individual on the gay guides data set. Essentially, gayguides$title[500] is a command telling R to search the gayguides data set for the occupation, name, and employer of the 500th person in the entire data set of 60,000 LGBTQ people.  
```

(@) The Year variable contains the year (numeric) of each entry. Can you find the earliest year in this dataset? (Using code, don't cheat and use the documentation!)
```{r}
 min(gayguides$Year)
[1] 1965
```


(@) How about the latest year? 
```{r}
 max(gayguides$Year)
[1] 1985
```


(@) Add another dataset from `DigitalMethodsData`. How many observations are included in this dataset? Run through some of the examples from above to learn about the dataset.
```{r}
> help(package="DigitalMethodsData")
> data(BostonWomenVoters)
> str(BostonWomenVoters)
'data.frame':	10702 obs. of  34 variables:
 $ Ward.Book                         : int  6 6 8 1 6 6 6 6 6 6 ...
 $ Volume                            : chr  "3" "102" "3" "3" ...
 $ Page.Number                       : chr  "3" "19" "17" "1" ...
 $ Date                              : chr  "8/20/20" "" "10/7/20" "8/19/20" ...
 $ Ward                              : int  6 NA 8 1 6 6 6 6 6 6 ...
 $ Precinct                          : chr  "5" "" "4" "7" ...
 $ Name                              : chr  "Josephine P. Helmer" "Theodosia E. Grice" "Anne M.T. Laurie" "Marion D. Todd" ...
 $ Street.of.Residence.on.April.1    : chr  "Tremont Street" "Shawmut Avenue" "Pinckney Street" "Monmouth Street" ...
 $ Street.Number.on.April.1          : chr  "626" "499" "52" "49" ...
 $ Unit                              : chr  "2" "2" "" "whole" ...
 $ Age                               : int  20 20 20 20 21 21 21 21 21 21 ...
 $ Country.of.Birth                  : chr  "France" "United States" "United States" "United States" ...
 $ State.or.Province.of.Birth        : chr  "" "Georgia" "New Hampshire" "Massachusetts" ...
 $ County.of.Birth                   : chr  "" "" "" "" ...
 $ Town.of.Birth                     : chr  "Cholet" "Darien" "Franklin" "Boston" ...
 $ Occupation                        : chr  "Housewife" "" "Clerk" "Stenographer" ...
 $ Place.of.Business                 : chr  "" "" "Federal Reserve Bank" "16 New Street" ...
 $ Court.of.Naturalization           : chr  "" "" "Single" "Single" ...
 $ Date.of.Naturalization            : chr  "" "" "" "" ...
 $ Husband.Name                      : chr  "William J. Helmer" "" "" "" ...
 $ Husband.Country.of.Birth          : chr  "United States" "" "" "" ...
 $ Husband.State.or.Province.of.Birth: chr  "Michigan" "" "" "" ...
 $ Husband.County.of.Birth           : chr  "" "" "" "" ...
 $ Husband.Town.of.Birth             : chr  "Pinconning" "" "" "" ...
 $ Husband.Date.of.Naturalization    : chr  "" "" "" "" ...
 $ Husband.Court.of.Naturalization   : chr  "" "" "" "" ...
 $ Father.Name                       : chr  "" "" "" "" ...
 $ Father.Country.of.Birth           : chr  "" "" "" "" ...
 $ Father.State.or.Province.of.Birth : chr  "" "" "" "" ...
 $ Father.County.of.Birth            : chr  "" "" "" "" ...
 $ Father.Town.of.Birth              : chr  "" "" "" "" ...
 $ Present.Residence                 : chr  "Same" "" "Same" "Same" ...
 $ Notes                             : chr  "" "Entry is partial and crossed out with note \"not of age until Jan 1921\"" "Note in Age field: \"Becomes 21 Oct. 28, 20\"" "Born: East Boston; Birthday: 1920-10-25; Note: \"Hold for Primary\"" ...
 $ X                                 : chr  "" "" "" "" ...
> head(BostonWomenVoters$title)
NULL
> head(BostonWomenVoters)
  Ward.Book Volume Page.Number     Date Ward Precinct                Name Street.of.Residence.on.April.1
1         6      3           3  8/20/20    6        5 Josephine P. Helmer                 Tremont Street
2         6    102          19            NA           Theodosia E. Grice                 Shawmut Avenue
3         8      3          17  10/7/20    8        4    Anne M.T. Laurie                Pinckney Street
4         1      3           1  8/19/20    1        7      Marion D. Todd                Monmouth Street
5         6     1A          10 10/13/20    6        1     Agnes P. Laffan                 Shawmut Avenue
6         6     1A          10 10/13/20    6        3    Augusta P. Davis                 Laconia Street
  Street.Number.on.April.1  Unit Age Country.of.Birth State.or.Province.of.Birth County.of.Birth Town.of.Birth
1                      626     2  20           France                                                   Cholet
2                      499     2  20    United States                    Georgia                        Darien
3                       52        20    United States              New Hampshire                      Franklin
4                       49 whole  20    United States              Massachusetts                        Boston
5                      177        21    United States              Massachusetts                        Boston
6                       19        21    United States              Massachusetts                        Boston
     Occupation    Place.of.Business Court.of.Naturalization Date.of.Naturalization      Husband.Name
1     Housewife                                                                     William J. Helmer
2                                                                                                    
3         Clerk Federal Reserve Bank                  Single                                         
4  Stenographer        16 New Street                  Single                                         
5 Drapery sewer                                                                                      
6   Book keeper                                                                                      
  Husband.Country.of.Birth Husband.State.or.Province.of.Birth Husband.County.of.Birth Husband.Town.of.Birth
1            United States                           Michigan                                    Pinconning
2                                                                                                          
3                                                                                                          
4                                                                                                          
5                                                                                                          
6                                                                                                          
  Husband.Date.of.Naturalization Husband.Court.of.Naturalization Father.Name Father.Country.of.Birth
1                                                                                                   
2                                                                                                   
3                                                                                                   
4                                                                                                   
5                                                                                                   
6                                                                                                   
  Father.State.or.Province.of.Birth Father.County.of.Birth Father.Town.of.Birth Present.Residence
1                                                                                            Same
2                                                                                                
3                                                                                            Same
4                                                                                            Same
5                                                                                            Same
6                                                                                            Same
                                                                   Notes X
1                                                                         
2 Entry is partial and crossed out with note "not of age until Jan 1921"  
3                            Note in Age field: "Becomes 21 Oct. 28, 20"  
4      Born: East Boston; Birthday: 1920-10-25; Note: "Hold for Primary"  
5                                                 Birthday on 1920-02-05  
6                                                                         
> BostonWomenVoters$title[500]
NULL
> BostonWomenVoters$title
NULL
> tail(BostonWomenVoters)
      Ward.Book Volume Page.Number     Date Ward Precinct                Name Street.of.Residence.on.April.1
10697        10    106          46  10/9/20   10        6 Sarah Clara Laforne                       K Street
10698        10    106          46 10/11/20   NA                A. Edna Emery              East Fifth Street
10699        10    106          47 10/11/20   NA                                          East Fourth Street
10700        10    106          48 10/11/20   10            Barbara M. Clancy                       L Street
10701        10    106          49 10/13/20   NA             E.D. DeVoy Ellen              East Sixth Street
10702        10     1A          14 11/24/20   10        9    Ellen T. Kennedy             East Fourth Street
      Street.Number.on.April.1 Unit Age Country.of.Birth State.or.Province.of.Birth County.of.Birth Town.of.Birth
10697                      340       NA                                                                          
10698                      533       NA                                                                          
10699                      707       NA                                                                          
10700                       53       NA                                                                          
10701                      668       NA                                                                          
10702                      835       NA                                                                          
      Occupation Place.of.Business Court.of.Naturalization Date.of.Naturalization Husband.Name Husband.Country.of.Birth
10697                                                                                                                  
10698                                                                                                                  
10699                                                                                                                  
10700                                                                                                                  
10701                                                                                                                  
10702                                                                                                                  
      Husband.State.or.Province.of.Birth Husband.County.of.Birth Husband.Town.of.Birth Husband.Date.of.Naturalization
10697                                                                                                                
10698                                                                                                                
10699                                                                                                                
10700                                                                                                                
10701                                                                                                                
10702                                                                                                                
      Husband.Court.of.Naturalization Father.Name Father.Country.of.Birth Father.State.or.Province.of.Birth
10697                                                                                                      
10698                                                                                                      
10699                                                                                                      
10700                                                                                                      
10701                                                                                                      
10702                                                                                                      
      Father.County.of.Birth Father.Town.of.Birth                       Present.Residence
10697                                                                                    
10698                                                                                    
10699                                                                                    
10700                                                                                    
10701                                                                                    
10702                                             Entry crossed out and "void" written in
                                            Notes X
10697 Entry crossed out, handwritten note: "void"  
10698 Entry crossed out; handwritten note: "void"  
10699                    Handwritten note: "void"  
10700                           Entry crossed out  
10701 Entry crossed out, handwritten note: "void"  
10702                                              
> head(BostonWomenVoters), n=16L
Error: unexpected ',' in "head(BostonWomenVoters),"
> head(BostonWomenVoters),n=16L
Error: unexpected ',' in "head(BostonWomenVoters),"
> head(BostonWomenVoters, n=16L)
   Ward.Book Volume Page.Number     Date Ward Precinct                Name Street.of.Residence.on.April.1
1          6      3           3  8/20/20    6        5 Josephine P. Helmer                 Tremont Street
2          6    102          19            NA           Theodosia E. Grice                 Shawmut Avenue
3          8      3          17  10/7/20    8        4    Anne M.T. Laurie                Pinckney Street
4          1      3           1  8/19/20    1        7      Marion D. Todd                Monmouth Street
5          6     1A          10 10/13/20    6        1     Agnes P. Laffan                 Shawmut Avenue
6          6     1A          10 10/13/20    6        3    Augusta P. Davis                 Laconia Street
7          6     1A          12 10/13/20    6        7      Alice E. Allen             East Newton Street
8          6     1A          12 10/13/20    6        1      Ella Silverman                 Emerald Street
9          6     1A          12 10/13/20    6        2    Sara A. Saccoach              Washington Street
10         6     1A          12 10/13/20    6        7       Edith W. Shaw             East Newton Street
11         6     1A          12 10/13/20    6        7       Helen D. Hart          East Brookline Street
12         6     1A          12 10/13/20    6        5   Gertrude R. Gorey                     Union Park
13         6     1A          14 10/13/20    6        7      Marion J. Port             East Newton Street
14         6     1A          14 10/13/20    6        6  Mildred Wedgeworth                  Malden Street
15         6      3           2  8/19/20    6        2   Matilda S. Cantor                    Rose Street
16         6      3           2  8/19/20    6        2         Mazie Welch                  Oswego Street
   Street.Number.on.April.1  Unit Age Country.of.Birth State.or.Province.of.Birth County.of.Birth Town.of.Birth
1                       626     2  20           France                                                   Cholet
2                       499     2  20    United States                    Georgia                        Darien
3                        52        20    United States              New Hampshire                      Franklin
4                        49 whole  20    United States              Massachusetts                        Boston
5                       177        21    United States              Massachusetts                        Boston
6                        19        21    United States              Massachusetts                        Boston
7                                  21    United States              Massachusetts                       Melrose
8                        65        21    United States              Massachusetts                        Boston
9                      1100        21    United States              Massachusetts                        Boston
10                                 21    United States                   New York                    Whitestone
11                       78        21    United States                 New Jersey                    Belleville
12                       46        21    United States              Massachusetts                      Brockton
13                       94        21    United States                      Maine                      Portland
14                      124        21    United States              Massachusetts                        Boston
15                       12     1  21    United States                Connecticut                   New Britain
16                       12     1  21    United States              Massachusetts                        Boston
       Occupation                 Place.of.Business Court.of.Naturalization Date.of.Naturalization      Husband.Name
1       Housewife                                                                                  William J. Helmer
2                                                                                                                   
3           Clerk              Federal Reserve Bank                  Single                                         
4    Stenographer                     16 New Street                  Single                                         
5   Drapery sewer                                                                                                   
6     Book keeper                                                                                                   
7      Bank clerk                                                                                                   
8       Saleslady            Superior Millinery Co.                                                                 
9         At home                                                                                                   
10          Clerk                                                                                                   
11 School teacher                                                                                                   
12 School teacher                                                                                                   
13        At home                                                                                     George F. Port
14   Stenographer                                                                                                   
15    Book keeper Auto car sales and [service?] Co.                  Single                                         
16      Housewife                                                                                      Patrick Welch
   Husband.Country.of.Birth Husband.State.or.Province.of.Birth Husband.County.of.Birth Husband.Town.of.Birth
1             United States                           Michigan                                    Pinconning
2                                                                                                           
3                                                                                                           
4                                                                                                           
5                                                                                                           
6                                                                                                           
7                                                                                                           
8                                                                                                           
9                                                                                                           
10                                                                                                          
11                                                                                                          
12                                                                                                          
13            United States                      Massachusetts                                        Boston
14                                                                                                          
15                                                                                                          
16            United States                      Massachusetts                                        Boston
   Husband.Date.of.Naturalization Husband.Court.of.Naturalization Father.Name Father.Country.of.Birth
1                                                                                                    
2                                                                                                    
3                                                                                                    
4                                                                                                    
5                                                                                                    
6                                                                                                    
7                                                                                                    
8                                                                                                    
9                                                                                                    
10                                                                                                   
11                                                                                                   
12                                                                                                   
13                                                                                                   
14                                                                                                   
15                                                                                                   
16                                                                                                   
   Father.State.or.Province.of.Birth Father.County.of.Birth Father.Town.of.Birth Present.Residence
1                                                                                             Same
2                                                                                                 
3                                                                                             Same
4                                                                                             Same
5                                                                                             Same
6                                                                                             Same
7                                                                                             Same
8                                                                                             Same
9                                                                                             Same
10                                                                                            Same
11                                                                                            Same
12                                                                                            Same
13                                                                                            Same
14                                                                                            Same
15                                                                                            Same
16                                                                               73 Compton Street
                                                                    Notes X
1                                                                          
2  Entry is partial and crossed out with note "not of age until Jan 1921"  
3                             Note in Age field: "Becomes 21 Oct. 28, 20"  
4       Born: East Boston; Birthday: 1920-10-25; Note: "Hold for Primary"  
5                                                  Birthday on 1920-02-05  
6                                                                          
7               Residence: Franklin Square House. Birthday on 1920-01-22.  
8                                                                          
9                                                  Birthday on 1920-01-27  
10              Residence: Franklin Square House. Birthday on 1920-05-18.  
11                                                Birthday on 1920-01-07   
12                                                 Birthday on 1920-10-04  
13                                                 Birthday on 1920-04-06  
14                                                 Birthday on 1920-05-16  
15                                                 Birthday on 1920-03-21  
16                                                 Birthday on 1920-05-06  
> min(BostonWomenVoters$Title)
[1] Inf
Warning message:
In min(BostonWomenVoters$Title) :
  no non-missing arguments to min; returning Inf
> min(BostonWomenVoters$Year)
[1] Inf
Warning message:
In min(BostonWomenVoters$Year) :
  no non-missing arguments to min; returning Inf
> min(BostonWomenVoters$Name)
[1] ""
> min(BostonWomenVoters$Place.of.Business)
[1] ""
> min(BostonWomenVoters$Occupation)
[1] ""
> min(BostonWomenVoters$Date)
[1] ""
> max(BostonWomenVoters$Volume)
[1] "V107"
> max(BostonWomenVoters$Street.of.Residence.on.April.1)
[1] "Yarmouth Street"
> min(BostonWomenVoters$Street.of.Residence.on.April.1)
[1] ""
> min(BostonWomenVoters$Precinct)
[1] ""
> max(BostonWomenVoters$Precinct)
[1] "9"
```

(@) What did you learn about the dataset? What can you tell me about it? 

>
```{r}
I learned that there are 10,702 observaions in BostonWomenVoters which is less than the gayguides data set. However, when I tried to find the minimum number of dates, occupations, volumes, places of businesses, and streets of residences, R came up with no return but when I attempted with max, it returned a value set. This was confusing because I was able the minimum for the gayguides but not BostonWomenVoters. However, the most fascinating part was when I observed the BostonWomenVoters was that despite most of the women living in Massachuesetts, there were some who were out of state as well.
```




CONGRATS! You've completed your first worksheet for Digital Methods II. That wasn't so bad, right? 

