---
title: "Worksheet 2: Data Structures"
name: David Askew
date: 2/7/2024
format: html
---

_This is the second in a series of worksheets for History 8510 at Clemson University. The goal of these worksheets is simple: practice, practice, practice. The worksheet introduces concepts and techniques and includes prompts for you to practice in this interactive document. When you are finished, you should change the author name (above), knit your document to a pdf, and upload it to canvas._

## Subsetting, Loops & Control Structures, and Functions

### Subsetting

Subsetting is the process of retrieving just the parts (a subset) of a large vector, list, or data frame. R has numerous operators for subsetting information or data structures. They allow you to perform complex operations with a small amount of code but can be hard to master because subsetting operators (`[[`, `[`, and `$`) interact differently with different vector types. 

We'll be talking about vectors throughout the first part of this worksheet and it might be useful to refresh our memories about what a vector is. 

> _A vector is a list of variables, and the simplest data structure in R. A vector consists of a collection of numbers, arithmetic expressions, logical values or character strings for example._

#### Selecting Multiple Elements

Lets start by creating a simple vector called x. 

```{r}
x <- c(15.4, 25, 2, 8.35, 5, 383, 10.2)
```

If you type `x` into your console, you will see that all of the values are printed out and preceded by a `[1]`. What the `[1]` refers to is that first number, 15.4. Its position in the list is 1. Each number in our list has an index number and that is how you retrieve specific positions in this vector. 

For example, we can use a positive integer to return elements at specific positions. Lets grab the value in the 3rd and 5th position.
```{r}
x[c(3,5)]
```
We can also use functions to manipulate our vector. Here we use `order()` to print the values contained in the vector in order. 
```{r}
x[order(x)]
```

Duplicate indices will return duplicate values. 
```{r}
x[c(3,3)]
```
(@) Create your own vector and return three values using the subsetting method above. 
```{r}
> y <- c(4, 8.1, 15, 16.2, 23, 42.3)
> y [c(2,7)]
[1] 8.1  NA
> x[c(2,4)]
[1]  8.1 16.2
> y[c(2,7)]
[1] 8.1  NA
> y[c(2,4)]
[1]  8.1 16.2
> y[order(y)]
[1]  4.0  8.1 15.0 16.2 23.0 42.3
> y[c(5,5)]
[1] 23 23
> y[(2,2)]
Error: unexpected ',' in "y[(2,"
> y[c(2,2)]
[1] 8.1 8.1
> y[c(6,6)]
[1] 42.3 42.3
```


Negative integers allow us to exclude values at certain positions. 
```{r}
x[-c(3, 1)]
```

(@) What happened here? Describe this in your own words. 

>
```{r}
What happened was that when I typed y[-c(4,2)], it excluded those numbers from the vector which got created. Therefore, instead of 4.0, 8.1, 15.0, 16.2, 23.0, and 42.3 all being displayed on the vector, it only showed 4.0, 15, 23, 42.3. Also, it will get rid of whatever position you type first before the comma and then the second one after the comma.

> y[-c(4,2)]
[1]  4.0 15.0 23.0 42.3
> y[-c(3,6)]
[1]  4.0  8.1 16.2 23.0
```

You can use either positive or negative integers to subset a vector but you **cannot mix the two**.

We can assign logical values to each value in a vector and use those to subset the data. Logical vectors select elements where the corresponding logical value is `TRUE`. Remember, we created a vector earlier and assigned it to x. Now, below, we assign logical values to each of the values in that vector. We're doing this by hand here, but you can imagine a scenario down the road where you use this technique to apply `TRUE` or `FALSE` values to a huge dataset dependent on some principal. When we run this, only the `TRUE` values are printed out. 
```{r}
#create a vector 
x[c(TRUE, FALSE, TRUE, TRUE, FALSE)]
```

We can also subset to get values that match a particular criteria. Below, each item in the vector is tested against this proposition and it returns the values that are `TRUE`.
```{r}
x[x > 7]
```

(@) What is going on in each of the lines above? Describe it in your own words. 

>
```{r}
> y[c(TRUE, TRUE, TRUE, TRUE, TRUE, TRUE)]
[1]  4.0  8.1 15.0 16.2 23.0 42.3
> y[c(TRUE, FALSE, TRUE, FALSE, TRUE, FALSE)]
[1]  4 15 23
> y[c(FALSE, TRUE, FALSE, TRUE, FALSE, TRUE)]
[1]  8.1 16.2 42.3
> y[y > 4]
[1]  8.1 15.0 16.2 23.0 42.3
> y[y > 16]
[1] 16.2 23.0 42.3
> y[y < 23]
[1]  4.0  8.1 15.0 16.2
> y[y < 16.2]
[1]  4.0  8.1 15.0

When I assigned TRUE to a certain number, it prints the number because it is TRUE, and thus shows it in the console. However, when I assigned FALSE to another number, it does not print since it is FALSE according to the console and therefore, does not show. Furthermore, when I used a greater than symbol for a number, it exclusively displays the rest which are greater than that value. The same thing happens with less than symbol which only displays numbers that are less than that value. 
```

Nothing returns the original vector. This isn't that useful here but can come in handy when you have more complex structures like matrices and arrays. 
```{r}
x[]
```

If your vector is named, you can also use character vectors to return elements with matching names. 
```{r}
(y <- setNames(x, letters[1:4]))
y[c("d")]
```
(@) What happened here? Explain this in your own words. 
```{r}
> (y <- setNames(x, letters[1:4]))
   a    b    c    d <NA> <NA> 
 4.0  8.1 15.0 16.2 23.0 42.3 
> y[c("d")]
   d 
16.2 
> (y <- setNames(x, letters[1:6]))
   a    b    c    d    e    f 
 4.0  8.1 15.0 16.2 23.0 42.3 
> y[c("b")]
  b 
  
What happened was that when I entered setNames for y in the console, it named those numbers and assigned a series of letters based on the ratio that I wanted to show in the console by typing [1:4] and [1:6]. When I typed 1:4, I was giving the console a command to print only the numbers which fell under the first four letters while not displaying the other two. This is similar to TRUE and FALSE where the machine will print the numbers that are TRUE based on the ratio and FALSE for all numbers which exceed that given ratio. ```

#### Matrices

We can subset higher dimensional structures, like matrices, too. First lets define a matrix. In R, **a matrix** is a collection of elements of the same data type (numeric, character, or logical) arranged into a fixed number of rows and columns. Since you are only working with rows and columns, a matrix is called two dimensional. To create a matrix you can use the `matrix()` function.
```{r}
matrix(1:9, byrow = TRUE, nrow = 3)
```

In the `matrix()` function:

  * The first argument is the collection of elements that R will arrange into the rows and columns of the matrix. Here, we use 1:9 which is a shortcut for c(1, 2, 3, 4, 5, 6, 7, 8, 9).
  * The argument byrow indicates that the matrix is filled by the rows. If we want the matrix to be filled by the columns, we just place byrow = FALSE.
  * The third argument nrow indicates that the matrix should have three rows.

(@) You try now, create a matrix that had five rows with numbers that range from 25 to 49. 
```{r}
> matrix (25:49, byrow = TRUE, nrow = 5)
     [,1] [,2] [,3] [,4] [,5]
[1,]   25   26   27   28   29
[2,]   30   31   32   33   34
[3,]   35   36   37   38   39
[4,]   40   41   42   43   44
[5,]   45   46   47   48   49
```

(@) Can you create a matrix that has 3 rows with only even numbers from 2 to 18? 
```{r}
> matrix <- seq (2, 18, by = 2)
> matrix (data = even_numbers, byrow = TRUE, nrow = 3)
     [,1] [,2] [,3]
[1,]    2    4    6
[2,]    8   10   12
[3,]   14   16   18

##### Subsetting Matricies 

The most common way to subset matrices and arrays is to use a simple generalization of the same subsetting method we used for vectors above. We can subset by supplying index for each dimension separated by a comma. So it looks like this: 
```{r}
a <- matrix(1:9, nrow=3)
colnames(a) <- c("A", "B", "C")
a
```
Here is our matrix. Its similar to the one we used above except I named each column using the `colnames()` function. 

So to subset this matrix and get just the first two rows I can do: 
```{r}
a[1:2,]
```
If I want to get the first two columns, I can do: 
```{r}
a[,1:2]
```
I could also exclude a column: 
```{r}
a[,-2]
```
Or get just one single value from the matrix: 
```{r}
a[2,3]
```
These skills are important because dataframes have the characteristics of both lists and matricies. 

So lets load the Gay guides data from our class pacakge. 
```{r}
library(DigitalMethodsData)
data("gayguides")
```

To subset this data we could pull two particular columns like this: 
```{r}
#gayguides[2:3]
```
That returns columns 2 and 3 (including every single row!). 

If we use two indices it'll behave more like a matrix. 
```{r}
gayguides[1:3,]
```

(@) Why is this different? What do we get compared to the previous example? 
```{r}
The reason why this is different is because since a matrix is two-dimensional, it requires two indices to treat them as two-dimensional objects which will return two colummns of data. When we use a ratio of 1:3 instead of 2:3, we print three of the total 60,698 observations which display in rows of three since nrow = 3 is demonstrated as :3. 
```


(@) Can you find just the city and state for the first 5 entries? 
```{r}
library(DigitalMethodsData)
> data("gayguides")
> #gayguides$city[1:5]
> gayguides$city[1:5]
[1] "Lake Placid"    "Troy"           "Salt Lake City" "Seattle"        "Troy"       
 gayguides$state[1:5]
[1] "NY" "NY" "UT" "WA" "NY"
```

(@) How about the street address and type for rows 2,555 to 2,560?

```{r}
>  gayguides$streetaddress[2555:2560]
[1] "721 Lincoln Ave."  "3231 N. Clark St." "721 Lincoln Ave."  "3231 N. Clark St." "721 Lincoln Ave."  "4427 W. Kennedy" 
> gayguides$type[2555:2560]
[1] "Restaurant,Bars/Clubs" "Bars/Clubs"            "Bars/Clubs,Restaurant" "Bars/Clubs"            "Bars/Clubs,Restaurant"
[6] "Bars/Clubs" 
```

(@) Load another dataset from `DigitalMethodsData`. Subset it to find some element of the data. Write the code below and then explain what element of the data you wanted to subset and why:
```{r}
`
> data("BostonWomenVoters")
> BostonWomenVoters$Precinct[1:100]
  [1] "5" ""  "4" "7" "1" "3" "7" "1" "2" "7" "7" "5" "7" "6" "2" "2" "6" "6" "9" "9" "8" "7" "7" "2" "6" "7" "9" "9" "7"
 [30] "7" "6" "7" "7" "6" "7" "5" "6" "6" "2" "7" "6" "7" "9" "7" "3" "7" "7" "7" "8" "6" "5" "8" "6" "7" "5" "6" "4" "8"
 [59] "9" "1" "7" "7" "7" "7" "8" "7" "6" "8" "8" "3" "5" "6" "7" "7" "4" "7" "8" "8" "5" "7" "3" "6" "4" "2" "3" "6" "9"
 [88] "6" "6" "8" "6" "9" "2" "6" "9" "7" "7" "8" "9" "9"
> BostonWomenVoters$Date[110:120]
 [1] "10/11/20" "10/11/20" "10/11/20" "10/13/20" "10/13/20" "10/13/20" "10/13/20" "10/13/20" "11/22/20" "11/23/20"
[11] "8/18/20"
 BostonWomenVoters$Name[2555:2570]
 [1] "Sophia Rudnick"       "Eliza S. Williams"    "Mary J. Welch"        "Julia Fitzpatrick"    "Anna J. Herb"        
 [6] "Olive M. Neilson"     "Josephine Anderson"   "Dora M. Horton"       "Anna E. Driscoll"     "Helen E. McIvor"     
[11] "Anna J. Ferreira"     "Grace A. McManus"     "Ursula A. Sears"      "Mary I. Kelly"        "Elizabeth M. Scheele"
[16] "Theresa M. Falabella"
> 
   I wanted to subset the precinct, dates, and names from the Boston Women Voters dataset. I started out by loading the data and typed in the ratio 1:100 to find the first 100 precincts that each of the women voted in for the election. By using this ratio, the program gave me the first 100 precincts of each voter. Then, I wanted to find the date of the 110th row to 120 for when each of these women voted. Afterward, I wanted to find women who voted further down from the list of the first 100 by typing ratio 2555:2570. This presented me with the 2555th woman all the way up to the 2570th that voted in the election. By using these different subsets and commands in R, I am able to navigate the dataset and load whatever data that I choose to print before typing the $ which then I can put a ratio that accesses specific data which I desire to print. The $ is especially useful in this endeavor. 

Another useful operator to know is `$` which we used a little bit in the previous worksheet. `$` is shorthand for accessing a variable in a dataframe. 
 
So for example both of these produce the same result. One is just easier to type. Because the output here is so large, I've commented these two lines out. Uncomment them and the run the block.
```{r R.options=list(max.print=10)}
#gayguides["title"]
#gayguides$title
```

Notice the above output is _huge_. We can combine `$` and `[` to drill down more specifically like we did earlier.
```{r}
gayguides$city[100:110]
```

(@) What does this code do? 

>
```{r}
It focuses on a specific quantity of data ranging from whatever range one wishes to print. The ratio after the $ followed by the data entry allows specific parts of the entire data set to be accessed. This is handy, especially when using data sets that have tens of thousands of observations and an individual wants to separate a specific part of that entire set.
```


We can also use `$` to create a new column in `gayguides`. 
```{r}
gayguides$mynewcolumn <- TRUE
gayguides$mynewcolumn[1:10]
```
What does this code do? It creates a new column in the gayguides data called `mynewcolumn` and it assigns `TRUE` to every row (all 60k). This is a bit of a unrealistic example. You probably won't want to add the same value to 60k+ rows but the concept here is useful and we'll use it later in this worksheet. 

### Control Structures: Loops & Choices

Loops and control structures can be indispensable tools when writing a script that does something for you. For example, maybe you want to write a script that goes out to a website, grabs a primary source, downloads it, and saves it to a file then repeats that hundreds of times until you have a set of primary sources. To write that kind of script you'd need loops and control structures to tell your program when to start, stop, and move to the next source. 

Loops and control structures are one of those programming patterns that appear in almost every language. The syntax and style change, but the core idea doesn't. This is a concept I'd recommend you master and it'll come in handy for years to come. 

**Choices** and **loops** are both known as types of **control structures.**

Choices, like `if`, allow you to run different code depending on the input. 

Loops, like `for` and `while` allow you to repeatedly run code, typically with changing options. 

Lets start with choices. 

#### Choices

The most important choice statement in R is an `if` statement.

An `if` statement allows you to check the condition of something and then determine what should be done based on that input. 

The basic format is this: 
```
if (condition) true_action 
#OR
if(condition) true_action else false action 
```

And it works like this
```{r}

if (2 < 3) print(TRUE)
if (3 < 2) print(TRUE)
```

The first example above works. Why? Because 2 is indeed less than 3. The second line has no output. Why? Becuase the conditions weren't met. 3 is not less than 2 and so it didn't print anything out.
But we could ask it to do something else instead:
```{r}
if (3 < 2) print(TRUE) else print("This is clearly false.")
```
Most often, if statements are more complex (known as compound statements) and so they use brackets like this: 
```{r}
x <- 71
if (x > 90) {
  "A"
} else if (x > 50) {
  "C"
} else {
  "F"
  }
```

(@) You are teaching History 100 for Clemson. Write an `if` statement that calculates whether the enrollment cap of 30 on your class has been met. 
```{r}
 students <- c("Marty", "Rick", "Fiona", "Lily", "Fred", "Doug", "Tyler", "Jack", "Nicholas", "Greg", "Phil", "Martin", "Kendall", "Logan", "Kailey", "Chris", "Carson", "Eliza", "Jennifer", "Benson", "Mordecai", "Margaret", "Dale", "Jade", "Joey", "Bill", "Tommy", "Tucker", "Joan", "Kennedy")
> if (length(students) == 30) {print("This class is full.")}
[1] "This class is full."
else if (length(students > 30)
> if (length(students) == 30) {print("This class is full.") } else if (length(students > 30){print("This class is over maximum capacity!")}
Error: unexpected '{' in "if (length(students) == 30) {print("This class is full.") } else if (length(students > 30){"
> if (num_students) > 30= {print("This class is full. No more students may enroll.")}
Error: unexpected '>' in "if (num_students) >"
if (Presidents < 6)print (TRUE)
```

(@) Create a list of presidents and store it in a variable. Use an `if` statement to check if the number of presidents in the list is more than 5. If its not indicate that too. 
```{r}
Presidents <- c("Abraham Lincoln", "Thomas Jefferson", "George Washington", "John Adams", "Theodore Roosevelt", "Franklin D. Roosevelt")
> 
n_Presidents <- 6
if n_Presidents > 5 {print("The number of Presidents exceeds 5")}
else {
```

(@) You can also use an if statement to check a variable in a dataframe. How might you use an if statement to check if the GayGuides dataset contains any year after 1990? (Hint: first try to figure out how to determine the latest year in the dataframe and then build an if statement that checks that and prints something out if its true or false. You should think about what kind of value in contained in the Year column, how to access it, and how to check for the latest value.) 

```{r}
if {(gayguides(year>$X1990)print(TRUE)}
```

(@) Reflect on the question above. How did you figure this out? What was the process you went through to build this chunk of code?
```{r}

```

> 

#### Loops

A **loop* is used to iterate over items in a vector. They typically follow this form:

```
for (item in vector) perform_action
```

For each item in the vector, perform_action is called once and the value of the item is updated each time. 

For example, 
```{r}

for (i in 1:3) {
  print(i)
}
```

What does this loop do? `1:3` is shorthand for 1, 2, 3. Run it in your terminal to see its output. This code says, for every number in the range 1:3, print the number.

We could do this with character vectors too. 
```{r}
presidents <- c("George Washington", "John Adams", "Thomas Jefferson")
for (p in 1:length(presidents)) {
  print(presidents[p])
}
```

(@) Why does this work? What does `length()` do and why do we use it on `presidents`?

> 
```{r}
Length() works because it is used to retrieve lengths of vectors and other objects. For 'presidents', it counts the number of presidents based on how many are in that assigned value and adjusts that object to a certain length. This is important to understand when applied to 'presidents' because we can determine how many values there are in that data set. 
```

Create a character vector that contains the title of each of your classes this semester. Can you write a `for` loop that prints out "I am enrolled in" and the name of each of your classes? ("I am enrolled in History 8150", "I am enrolled in History 8000".....etc). Hint: you'll need a function to combine the output and some text of your choosing inside the function. 

```{r}

```

Sometimes we want our loop to terminate itself early. There are two statements that help us do that: 

* `next` exits the current iteration
* `break` exits the entire for loop

```{r}
for (i in 1:10) {
  if (i < 3)
    next
  
  print (i)
}
```

`Next` skips values until the criteria is met. So in the above example it skips numbers 1 and 2 because they don't match the criteria. But in some cases we may not want to skip the entries but rather exit the loop entirely. Something like this: 
```{r}
for (i in 1:10) {
  if (i < 3)
    next
  
  if (i >= 5)
    break
  
  print (i)
}
```

(@) What happened here? Why did the loop only print 3 and 4? 


(@) In the state population data, can you write a loop that pulls only states with populations between 200,000 and 250,000 in 1800? You could accomplish this a few different ways. Try to use next and break statements. 
```{r}
library(DigitalMethodsData)
data("statepopulations")
for (i in 1:length(statepopulations$X1800)){

   if (is.na(statepopulations$X1800[i])){
    next
  } else if (statepopulations$X1800[i] < 200000){
    next
  } else if (statepopulations$X1800[i] > 250000){
    next
  }
    print(statepopulations$STATE[i])
  }
```

(@) What if we wanted to loop through the gay guides data? Can you write a loop that iterates through the cities in the Gay Guides data and returns the title of any location in Greenville? 
```{r}
for($city(gayguides$city)){location <- gayguides[gayguides$city == city,]}
title <- location 
```

`while` is another useful tool for writing loops. While will run a loop _while_ a condition is true. Once that condition is not true, it exits. For example: 
```{r}
i <- 1
while (i < 6) {
  print(i)
  i <- i+1
}
```
Its important to note here that `i` is defined outside the loop. Its initial value is 1 but each time the loop runs it gets 1 added to it. So the first time it runs, i = 1 and the loop prints the value then adds 1 to i. Now i = 2 but since that is still less than 6 the loop continues and prints the value then adds 1 to i. Eventually, i=6 and since 6 is not less than 6 the loop exits without printing. 

`While` doesn't get used as often in R because there is often a more efficient way to do the same thing. But in other languages `while` gets used much more often. Its a good pattern to be familiar with even if you don't use in frequently in R. 

## Functions 

A **function** is a set of statements that are organized together to perform a specific task. You've already been using functions because R has a number of built in functions that are available to users. For example. `print()` is a function. 

To be more specific, a **function** is a set of code that performs a task and returns a result. As you advance your programming skills you'll probably have certain tasks that you perform frequently. Once you've run a chunk of code several times its good practice to turn it into a function so it can be repeatedly used. 

A function in R takes any arguments that may be necessary to perform the actions contained within the function, performs those functions, and then returns any result in the console. 

For example, the below function takes a number as an input, multiplies it by 5 and then returns the result. 
```{r}
myfunction <- function(y){
  myvalue <- y * 5
  print(myvalue)
}
```
When you run this nothing is returned. Why? You've loaded the function but you haven't called it yet. We can do that like this: 
```{r}
myfunction(5)
```

You'll notice that the variable we created inside the function, `myvalue`, doesn't show up. Unless we write code asking R to return the value of that variable, it'll run invisibly in the background and only give us back the result we ask for. This comes in handy when you are writing more complex functions and need to store bits of data temporarily for a calculation that is being run by the function. 

Here's another example: 
```{r}
historyStudents <- function(studentname, class) {
  statement <- paste("Hello ", studentname, ". Welcome to ", class, "!", sept="")
  statement
}
```
(@) Can you run this function? 
```{r}
historyStudents <- function(studentname, class) 
{statement <- paste ("Hello " , "studentname", ". Welcome to" , class, "!", sep="")statement}
historyStudents <- function(studentname, class)
+ {statement <- paste ("Hello " , "student", ". Welcome to" , class, "!" sep="")statement}
Error: unexpected symbol in:
"historyStudents <- function(studentname, class)
{statement <- paste ("Hello " , "student", ". Welcome to" , class, "!" sep"
> {statement <- paste ("Hello " , "studentname", ". Welcome to" , class, "!", sep="")statement}
Error: unexpected symbol in "{statement <- paste ("Hello " , "studentname", ". Welcome to" , class, "!", sep="")statement"
> 
```


There are several components to a function. 

* **Function Name**:This is the actual name of the function. It is stored in R environment as an object with this name. In the above example, the function name is `historyStudents`
* **Arguments**: An argument is a placeholder. When a function is invoked, you pass a value to the argument. Arguments are optional; that is, a function may contain no arguments. Also arguments can have default values. In the example above, the function takes two arguments: a student's name and the name of a class. 
* **Function Body**: The function body contains a collection of statements that defines what the function does. In the above example the function body is the code inside the `{` and `}` brackets. 
* **Return Value**: The return value of a function is the last expression in the function body to be evaluated. In the above example, our return value is `statement` which prints our welcome statement using the names we provided in the arguments. 

(@) Write a function that takes a string of text and counts the number of characters. The function should return "There are xx characters in that string." 
```{r}
male <- function(x)
if (x > 7){
  "tall"
}else if (x > 15) {}
  "short"
}
```

(@) Reflect on the process of building the above function. How did you go about figuring this out? 

>

The body of a functions can use any of the techniques we've learned in this worksheet like loops and if statements. For example: 
```{r}
grade <- function(x) {
  if (x > 90) {
  "A"
} else if (x > 50) {
  "C"
} else {
  "F"
  }
}


grade(85)
```
We could run this function as many times as we want. 
```{r}
grade(95)
grade(75)
```

(@) In the example below, why doesn't George Washington's grade get printed out? 
```{r}
GeorgeWashington <- grade(60)

The reason why is because it is missing the print() function. Therefore, R may assign a value of 60 but will not print or display it in the console. Therefore, you must add print(GeorgeWashington).

GeorgeWashington <- grade(60)
print(GeorgeWashington)
```

>
```{r}


Here's a more complex example: 
```{r}
gg.locations <- function (city, year) {
  data(gayguides)

  for (i in 1:length(gayguides$city)) {
    if (gayguides$city[i] == city && gayguides$Year[i] == year) {
     print(paste("Found a location called ", gayguides$title[i]))
    } else {
      next 
    } 
  }
}
gglocations(Greenville, SC)
```

(@) Write a function that uses the things you've learned in this worksheet. The function should accept two arguments, city and year. It should pull all the locations in that city and year and return a statement that says "Found a location called xxx". (Where x is the title of a location in that year and city.) 
```{r}
Location <- function(Oslo, 2022) {print("Found a location called Norway")}
```

(@) Use the Boston Women Voter dataset. Write a function that accepts the name of an occupation. Your function should return new dataframe that includes the records for every voter listed with that occupation.  
```{r}
bwv.occupation <- function(occupation){
  require(DigitalMethodsData)
  data("BostonWomenVoters")
  for (i in 1:length(BostonWomenVoters$Occupation)){
  if (BostonWomenVoters$Occupation[i] == occupation){
    print(BostonWomenVoters$Occupation[i])
    }
    
  }
  
}
    

bwv.occupation("Student")
bwv.occupation("Housewife")
```

